---
title: "Using stochastic parameters"
date: "Last updated: 15 April 2020"
vignette: >
  %\VignetteIndexEntry{Using stochastic parameters}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
version_date <- lubridate::ymd("2020-04-12")
knitr::opts_chunk$set(echo = TRUE, cache=FALSE, eval=TRUE, collapse = TRUE,
                      tidy.opts=list(width.cutoff=60), tidy=FALSE)
```


```{r model setup}
library(sirplus)
library(ggplot2)

s.num <- 1000  # number susceptible
i.num <- 15  # number infected
q.num <- 5  # number in self-isolation
h.num <- 1  # number in the hospital
nsteps <- 101 # number of steps (e.g. days) to simulate
nsims  <- 20

control <- control_seiqhrf(nsteps = nsteps)
init <- init_seiqhrf(s.num = s.num, i.num = i.num, q.num = q.num, h.num = h.num)
```

Instead of running all `nsims` simulations with fixed parameter values, 
sirplus enables the `act.rate.*` and `inf.prob.*` (where `*` is one of `i`, `e` and `q`)
to be randomly generated from a probability distribution, 
which can be seen as a prior. 

## 1. Using fixed parameter values across time with default prior distributions

We use function `select_prior` to specify priors.
If default prior distributions are desired, we only need to input a vector of the names of parameter that are to be generated stochastically. 

If not specified, the default values are as follows:

| Variables  | Distribution     | Parameters (of the prior distribution) and default values    |
|-------------|----------------------------| --------------------------------------------------------|
| **act.rate.\* **     | Gaussian  | `mean = param$act.rate.*` and `sd = 0.1`  | 
| **inf.prob.\* **     | Beta  | `a = 50` and `b` is calculated s.t. the expectation equals `param$inf.prob.*` | 

```{r fixed prior}
param <- param_seiqhrf()
priors0 <- select_prior(param, c("act.rate.q", "act.rate.i", "inf.prob.i", "inf.prob.e"))
priors0
```

We can see how the simulation is different from the baseline simulation which uses constant parameters. 
```{r fixed sim}
baseline_sim <- seiqhrf(init, control, param)
baseline_plot <- plot(baseline_sim, start_date = lubridate::ymd("2020-01-01"), comp_remove = c('s.num', 'r.num'), plot_title = 'Baseline Model')

sim0 <- seiqhrf(init, control, param, priors0)
sim0_plot <- plot(sim0, start_date = lubridate::ymd("2020-01-01"), comp_remove = c('s.num', 'r.num'), plot_title = 'Default priors')

gridExtra::grid.arrange(baseline_plot, sim0_plot)
```



## 2. Using time varying prior distributions with default prior distributions

If a parameter in `param` is a vector (i.e. its value varies over time), its prior distribution will be different across time points as well. 

```{r vary prior}
param <- param_seiqhrf(act.rate.i = seq(10, 5, -0.05), inf.prob.i = seq(0.1, 0.05, -0.0005))
priors1 <- select_prior(param, c("act.rate.q", "act.rate.i", "inf.prob.i", "inf.prob.e"))
priors1
```

We can plot the priors and see how they change with time.
If `nstep` is larger than 10, the prior curves of only 10 (evenly distributed) time points will be plotted (see the Step label). 
```{r vary plot, warning=FALSE}
gg1 <- plot(priors1)
gridExtra::grid.arrange(gg1[[1]], gg1[[2]], gg1[[3]], gg1[[4]])


sim1 <- seiqhrf(init, control, param, priors1)
plot(sim1, start_date = lubridate::ymd("2020-01-01"), comp_remove = c('s.num', 'r.num'), plot_title = 'Priors with larger variance')
```

## 3. Using time varying parameter values with the same prior distribution 

Another alternative to incorporate time varying parameters, is to simulate only for the first time point, 
parameters used for the latter time points are shifted from the input values in `param` by the same amount as the first time point. 
By doing so, we preserve the desired trend/change of the parameter across time by avoiding excessive variance,  while allowing some randomness in the intercept. This can by achieved by simply setting `shift_only == TRUE`.

```{r shift_only prior}
priors2 <- select_prior(param, c("act.rate.q", "act.rate.i", "inf.prob.i", "inf.prob.e"), shift_only = TRUE)
sim2 <- seiqhrf(init, control, param, priors2)
```

Take `act.rate.i` for example, the increment between time points are preserved, but for each simulation the starting point is randomly generated. 
```{r shift_only pars, echo=FALSE}
cat("Generated act.rate.i: \n")
tmp <- sim2$sampled.pars[["act.rate.i"]][1:3, 1:3]
rownames(tmp) <- c("sim1", "sim2", "sim3")
colnames(tmp) <- c("t = 1", "t = 2", "t = 3")
round(tmp, 3)

tmp <- param[["act.rate.i"]][1:3]
names(tmp) <- c("t = 1", "t = 2", "t = 3")
cat("\n\nOriginal input of act.rate.i in param: \n")
tmp
```

## 4. Using user specified prior distributions 

We can also manually specify the parameters in the prior distributions. 
For example, the following parameter setting results in distributions with larger variance. 
```{r}
priors3 <- select_prior(param, c("act.rate.q", "act.rate.i", "inf.prob.i", "inf.prob.e"), prior.dist = c("gaussian", "gaussian", "beta", "beta"),
                       prior.pars = list(list(sd = 1), list(sd = 2), list(a = 1), list(a = 10)))
priors3
gg3 <- plot(priors3)
gridExtra::grid.arrange(gg3[[1]], gg3[[2]], gg3[[3]], gg3[[4]])
```

As expected, this leads to a larger variance in simulations as well:
```{r, warning=FALSE, message=FALSE}
sim3 <- seiqhrf(init, control, param, priors3)
plot(sim3, start_date = lubridate::ymd("2020-01-01"), comp_remove = c('s.num', 'r.num'), plot_title = 'Default priors')
```
